---
title: "ToothGrowth Inferential Data Analysis"
author: "Charles Yoo"
date: "June 19, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Overview

This project investigages the correlation that various dosages and supplements have on tooth length using the ToothGrowth dataset.  The null hypothesis is that OJ has a stronger efficacy on tooth growth than Vitamin C.

## Load Data

```{r eval = TRUE}
data(ToothGrowth)
```

## Summary of Data
```{r eval = TRUE}
# description of variables
str(ToothGrowth)

# summary of dataset
summary(ToothGrowth)

# look at first 6 rows
head(ToothGrowth)

# count number of rows
nrow(ToothGrowth)
dim(ToothGrowth)

# exploration of data
unique(ToothGrowth$dose)
unique(ToothGrowth$supp)

# table comparing dose to supp
table(ToothGrowth$dose, ToothGrowth$supp)
```

## Comparison of dose to supp

```{r eval = TRUE}
ggplot(ToothGrowth, aes(x = supp, y = len)) +
        geom_boxplot(aes(fill = factor(supp))) +
        facet_grid(. ~ dose) +
        xlab("Supplement") +
        ylab("Tooth Length") +
        ggtitle("Toothgrowth length vs supplement by dose")
```

We see a visual correlation between dose size and tooth length.  The higher the dosage of the supplement affects a longer tooth length.  We also see that the type of supplement may have an effect on tooth length as well at the lower dosages.

## Confidence intervals

*Null-hypotheis*

OJ supplement has a more impact than Vitamin C.

#### T-test supplment as factor

Assign variables
```{r eval = TRUE}
length <- ToothGrowth$len
supplement <- ToothGrowth$supp
dosage <- ToothGrowth$dose
```

T-test on supplement
```{r eval = TRUE}
t.test(length[supplement == "OJ"], length[supplement == "VC"], paired = FALSE, var.equal = FALSE)
```
The confidence interval [-0.1710156, 7.5710156] and the p value 0.06063 shows a significance. The difference in mean is not significant enough to support the null hypothesis.

#### T-test dosages as factor

Assign variables
```{r eval = TRUE}
dose0510 <- subset(ToothGrowth, dose %in% c(0.5, 1.0))
dose0520 <- subset(ToothGrowth, dose %in% c(0.5, 2.0))
dose1020 <- subset(ToothGrowth, dose %in% c(1.0, 2.0))
```

T-test on doses
0.5 & 1.0
```{r eval = TRUE}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = dose0510)
```
The confidence interval [-11.983781, -6.276219] and the p-value 1.268e-07, essentially 0, shows a significant correlation between tooth length and dose levels.  There is a significant difference between the dosages that the null hypothesis is true.

0.5 & 2.0
```{r eval = TRUE}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = dose0520)
```
The confidence interval [-18.15617, -12.83383] and the p-value 4.398e-14, essentially 0, shows a significant correlation between tooth length and dose levels.  There is a significant difference between the dosages that the null hypothesis is true.

1.0 & 2.0
```{r eval = TRUE}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, data = dose1020)
```
The confidence interval [-8.996481, -3.733519] and the p-value 1.906e-05, essentially 0, shows a significant correlation between tooth length and dose levels.  There is a significant difference between the dosages that the null hypothesis is true.

#### T-test supplement as factor within dosage levels

Assign variables
```{r eval = TRUE}
dose05 <- subset(ToothGrowth, dose == 0.5)
dose10 <- subset(ToothGrowth, dose == 1.0)
dose20 <- subset(ToothGrowth, dose = 2.0)
```

T-test on supplement with doses
0.5
```{r eval = TRUE}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = dose05)
```
The confidence interval [1.719057, 8.780943] and the p-value 0.006359, essentially 0, shows a significant correlation between tooth length and dose levels.  There is a significant difference between the dosages that the null hypothesis is true.

1.0
```{r eval = TRUE}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = dose10)
```
The confidence interval [2.802148, 9.057852] and the p-value 0.001038, essentially 0, shows a significant correlation between tooth length and dose levels.  There is a significant difference between the dosages that the null hypothesis is true.

2.0
```{r eval = TRUE}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, data = dose20)
```
The confidence interval [-0.1710156  7.5710156] and the p-value 0.06063, essentially 0, shows a significant correlation between tooth length and dose levels.  The difference in mean is not significant enough to support the null hypothesis.

## Conclusions

* There is no significance difference in tooth lengths when looking at supplements only.  
* There is a significance difference in tooth lengths when looking at dosages only.  

### References

http://rstudio-pubs-static.s3.amazonaws.com/56907_9a9016c4f4df4d10aa7afaed5eed1d77.html  
https://rpubs.com/YevgenyY/si_cp2  
http://rstudio-pubs-static.s3.amazonaws.com/26135_b10068633b5948848d2c28fc0a752d78.html  
